Artificial Neural Network based critical conditions for Dynamic Recrystallization of Medium Carbon Steel and Application

### 2023-08-25 Article.tex
Update Section 2

%----------------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%----------------------------------------------------------------------------------
Dynamic recrystallization is a process observed in specific materials, mainly metals and alloys, when undergoing non-reversible plastic deformation.
It is a phenomenon frequently seen in hot working operations, including hot rolling, hot forging, and hot extrusion.
When a metal or alloy undergoes significant plastic deformation at high temperatures, its existing grain structure is disrupted.
During this deformation process, new, smaller grains are formed, which is referred to as dynamic recrystallization (DRX).
DRX differs from static recrystallization (SRX), which takes longer to form new grains and occurs at lower temperatures.
The amount of energy stored in the metal due to plastic deformation is the driving force behind the DRX process.
As the metal deforms, internal energy increases.
At high temperatures, metal atoms have enough mobility to rearrange themselves, forming new equiaxed grains free of strain.
Typically, these newly formed grains are much smaller than the original grains of the undeformed material.
Their growth is accelerated by the stored energy and high temperatures.
DRX provides several benefits in the hot forming process by refining the grain size, enhancing the mechanical properties of the material, including strength and toughness.
Additionally, it decreases the total strain energy, resulting in a more uniform microstructure and better material properties.

During high-rate hot deformation, dislocation density typically rises while a dynamic recovery (DRC) helps to mitigate such increases.
DRX occurs when the dislocation density rises to an extent beyond the DRC’s accommodation.
This is especially true for low and medium stacking fault energies \OP{what is it?} in fcc metals.\OPP{can you explain what you mean?}
Therefore, understanding the critical conditions for dynamic recrystallization (DRX) is important for modeling industrial processes.
However, the extent of DRX is influenced by factors such as material composition, deformation temperature, strain rate, and degree of deformation.
DRX is a crucial phenomenon that must be considered and controlled in industrial processes that use hot working to shape metals and alloys into products.
By optimizing the parameters, engineers can attain the desired microstructures and mechanical properties in the final product.
Various techniques have been proposed by researchers to establish DRX's critical conditions such as metallography.
However, this technique demands abundant sampling prior to and after the critical deformation.
Additionally, the cooling phase entails phase changes from the hot working temperature, which alter the deformed structure, rendering the metallographic analysis complex.
Hence, simpler techniques such as the development of analytical models are required to determine DRX initiation.
In this regard, multiple studies have been carried out aiming to analytically identify the initiation of DRX.
For instance, Barnett \eal \cite{barnett2000predicting} determined the critical strain for the initiation of DRX using the kinetics of static recrystallization (SRX), which was adapted to allow SRX to begin before the end of deformation.
This approach defines the critical stress required for the initiation of SRX during deformation.
According to Gottstein \eal \cite{gottstein2004prediction}, the critical strain for initiating DRX can also be predicted based on a work hardening model of dislocation density.
The onset of dynamic recrystallization (DRX) may be identified through the inflection point in the strain hardening rate $\theta=\frac{\partial \sigma}{\partial \varepsilon}$ where $\sigma$ is the stress and $\varepsilon$ is the strain.
This understanding is crucial for materials processing and design purposes.

The DRX phenomenon, based on the inflection point, was first studied by Ryan and McQueen \cite{ryan1989dynamic, ryan1990dynamic, ryan1990flow}.
They observed the appearance of an inflection point on the $\theta=f(\sigma)$ \OP{I suggest the following writing: $\theta(\sigma)$ instead of $\theta=f(\sigma)$} curve before the peak stress $\sigma_p$ and attributed the beginning of DRX and the critical conditions to this point.
Later, Poliak and Jonas \cite{Poliak-1996,ei2003initiation,ei2003critical,jonas2003critical} confirmed the hypothesis of DRX describing the inflection point by demonstrating that it is associated with thermodynamic energy released during dislocation movements.
This indicates the onset of an additional softening mechanism, in addition to DRC.
Najafizadeh and Jonas \cite{najafizadeh2006predicting} simplified Poliak and Jonas' thermodynamic model using an analytical technique to determine the critical conditions for initiating DRX.
Their widely-used approach relies on a third-degree polynomial to describe the curve $\theta=f(\sigma)$ and the second derivative to identify the critical stress $\sigma_c$ and strain $\varepsilon_c$.
Once calculated, the critical conditions serve as input data for the Johnson-Mehl-Avrami-Kolmogorov (JMAK) model \cite{Avrami-1939}.
This model is commonly utilized to estimate the volumetric fraction of dynamic recrystallization $X_{drx}$ during hot working.
Li \eal \cite{li2015experimental} investigated the DRX properties of micro-alloyed plastic molding steel by utilizing the Avrami kinetics model equation and the Estrin and Mecking mathematical model \cite{estrin1984unified,mecking1981kinetics} to determine essential parameters.

@@

The precision of the critical DRX conditions is closely linked to the method of obtaining the curve $\theta=f(\sigma)$.
Using experimental raw data directly does not yield functional curves, thus smoothing the $\sigma=f(\varepsilon)$ \OP{I suggest the following writing: $\sigma(\varepsilon)$ instead of $\sigma=f(\varepsilon)$} curve as much as possible or approximating it with a polynomial function that can reproduce the curve and deducing the $\theta=f(\sigma)$ curve from it is commonly used.
The drawback of both of these methods is that they do not always ensure a satisfactory smoothing of the curve $\sigma=f(\varepsilon)$.
Nevertheless, the Universal Approximator, Artificial Neural Networks (ANN), has demonstrated effectiveness in various domains, such as predicting the flow behavior of materials.
Hence, this study offers a replacement of the two smoothing methods with an ANN to procure the critical conditions for DRX.
In other words, the proposed method suggests utilizing an artificial neural network (ANN) to directly model the relationship between stress $\sigma$ and rate hardening $\theta$, removing the need for explicit smoothing of experimental data.
By properly training the ANN with relevant data, it can learn the underlying relationship between stress and dynamic recrystallization (DRX), ultimately providing a more precise estimation of the critical conditions for DRX.

@@

The experimental tests on which this work is based are strictly identical to those already published by Tize Mha \eal \cite{TizeMha-2023}.
For further information, readers may refer to this other publication by our research group.
We will, however, list here-after the main points required to understand the proposed approach.
The first publication focused on developing a constitutive law capable of predicting material flow, for which a comparative study between 5 analytical models and a neural network-based model was conducted.
At the conclusion of this comparative study, the Artificial Neural Network (ANN) model proved to be the best performing.
In this new study, the ANN model will be utilized for predicting dynamic recrystallization, as explained in the introduction.
It's worth noting that the ANN model identified in the first publication will also be employed for managing the plastic behavior of the material.

%----------------------------------------------------------------------------------
\subsection{Experimental procedure and compression tests results\label{subsec:ExperimentalProcedure}}
%----------------------------------------------------------------------------------
The material used for this study is a medium-carbon steel with the chemical composition shown in Table \ref{tab:Composition}.
\begin{table}[H]
\centering
\caption{Chemical composition of medium carbon steel. Fe = balance.}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{LCCCCCCC}
\hline
\textbf{Element} & \textbf{C} & \textbf{Mn} & \textbf{Mo} & \textbf{Si} & \textbf{Ni} & \textbf{Cr} & \textbf{Cu} \\
\hline
\textbf{Wt~\%} & $0.30$ & $0.89$ & $0.52$ & $0.34$ & $0.68$ & $1.86$ & $0.17$ \\
\hline
\end{tabularx}
\label{tab:Composition}
\end{table}
Hot compression tests of cylinders with an initial diameter of $\phi=10$~mm and a height of $h=15$~mm were carried out on a Gleeble-3800 thermomechanical simulator, for $5$ temperature levels (between $1050\celsius$ and $1250\celsius$), and $6$ strain rates (between $0.001~\ps$ and $5~\ps$).
The imposed displacement is set to $d=6$~mm, leading to a reduction of $40\%$ of the initial height.
To minimize friction during testing, thin tantalum sheets were used as a lubricant on the contact surface of the anvils and specimens.
To eliminate thermal gradients, the samples were heated at a rate of $2$\celsius/s to a temperature of $1260\celsius$ and held at that temperature for $5$~min.
They were then cooled to the test temperature at a rate of $1$\celsius/s and then held at constant temperature for $1$~min prior to forming.
During compression, the specimen temperature is kept constant by the machine's thermal control system.
The sample is then rapidly cooled to freeze the microstructure for later analysis.

The raw stress-strain data are exported from the Gleeble thermomechanical simulator as true stress and true strain, and the set of $30$ flow stress curves $\sigma$ versus strain $\varepsilon$ obtained from compression tests for each test condition is reported in Figure \ref{fig:RawData}.
All strain/stress data is composed of $701$ equidistant strain values from $\varepsilon=0.0$ to $\varepsilon=0.7$ in $0.01$ increments.
\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/rawData}
\caption{Stress--strain curves of medium carbon alloy extracted from the Gleeble device for 5 temperatures ($T$) and 6 strain rates ($\mdot\varepsilon$).}
\label{fig:RawData}
\end{figure}

The flow stress ($\sigma$) increases with increasing strain rate ($\mdot\varepsilon$) but decreases with increasing temperature ($T$).
It is noteworthy that the flow stress is also influenced by the strain.
For the lowest strain rates, the flow stress increases up-to $\sigma_p$ with the strain until reaching a value of approximately $\varepsilon_p=0.2$ to $0.3$.
Subsequently, it decreases to maintain a relatively constant value throughout the test.
Conversely, for the highest strain rates (above $1~\ps$), the flow stress consistently increases during the entire test duration.
The slight increase in stress at low strain rates with large strain values is due to friction between the sample and the anvil during testing as reported by Galos \eal \cite{Galos-2022}.
This frictional effect becomes more pronounced as lubrication decreases over time.

The initial stress increase observed during deformation up to $\varepsilon=0.1$ is attributed to work hardening (WH).
Subsequently, from $\varepsilon=0.1$ to $0.2$, the flow stress demonstrates a continuous decline as the stress level increases until a peak or inflection point is reached, indicating the dominant influence of thermal softening surpassing work hardening.
At this stage, the stress-strain curve exhibits three different patterns as the strain increases: (i) gradual decrease to a steady state with DRV/DRX softening.
This pattern is observed for all deformation temperatures and strain rates ranging from $\mdot\varepsilon=0.001~\ps$ to $0.1~\ps$, except for those at $1050\celsius$ and $1100\celsius$; (ii) higher stress levels without significant softening and work hardening at $1050\celsius$ and $1100\celsius$ with a strain rate of $0.1~\ps$; (iii) continuous increase with significant work hardening observed for all deformation temperatures and a strain rate of $1~\ps$.
This suggests that softening due to DRX occurs at high temperatures and low strain rates.
In contrast, at higher strain rates and lower temperatures, the rate of softening due to DRX is slowed down by higher work hardening rates.
As a result, both the peak stress $\sigma_p$ and the onset of steady-state flow occur at higher strain levels.
The drop in stress observed at all temperatures and strain rates of $\mdot\varepsilon=0.001-5.0~\ps$ is due to the occurrence of DRX.

DRX is an important phenomenon for controlling the microstructural and mechanical properties of a material during hot working.
In some materials, such as aluminum, DRX can balance work hardening, resulting in the rapid appearance of a plateau.
However, in many austenitic steels, the kinetics of restoration are weak and, as a result, DRX can be triggered at a critical condition of strain accumulation.
Because of the significant influence of DRX on the material's flow stress at high temperature, this inevitably affects the material's microstructure after hot working, and prediction of the critical conditions for the onset of DRX is therefore crucial to good material characterization.
Several factors, such as the chemical composition of the material, the grain size before deformation, the mode of deformation and the deformation parameters (deformation rate, temperature and deformation) condition DRX, which can only be triggered if the critical deformation is reached, and this deformation can be determined using metallography.
However, this technique requires extensive sampling before and after critical deformation.
In addition, phase changes during cooling from the hot working temperature modify the deformed structure, making metallographic analysis difficult.
Hence the need for simpler techniques, such as the development of analytical models.
With this in mind, a number of studies have been carried out to determine analytically the initiation of DRX.

In this paper we will use the Johnson-Mehl-Avrami-Kohnogorov (JMAK) phenomenological model \cite{Avrami-1939} to compute the evolution of DRX of the material during the deformation.
The raw data acquired from the Gleeble thermosimulator will subsequently be used to identify the material flow law and the parameters of the DRX model.
The parameters of JMAK model are obtained through a two-step process.
Firstly, the critical stresses and strains are determined, where the ANN model is utilized to obtain well-filtered curves.
Secondly, these critical values are employed as input data for the calculation of parameters in the JMAK model, aiming to predict the volume fraction of dynamic recrystallization.

@@

In this Section we will determine the critical deformation required to initiate recrystallization.
Ryan and McQueen \cite{Ryan-1989, Ryan-1990, Ryan-1990-2} were the first to study this phenomenon, observing the appearance of an inflection point on the curve $\theta(\sigma)=\frac{\partial \sigma}{\partial \varepsilon}$ before the peak stress $\sigma_p$ and attributed to this inflection point the start of DRX and therefore the critical values $\varepsilon_c$ and $\sigma_c$ of the strain and the stress respectively.
Later Poliak and Jonas \cite{Poliak-1996, Poliak-2003, Poliak-2003-2, Jonas-2003} further demonstrated that this inflection point is associated with a thermodynamic free energy released during dislocation motion, and consequently endorse the DRX hypothesis describing this inflection point.
Najafizadeh and Jonas \cite{Najafizadeh-2006} simplified the thermodynamic model of Poliak and Jonas by developing an analytical technique, widely used today, to determine the critical conditions for the initiation of DRX.
In their approach, a polynomial of $3^\text{th}$ degree is adopted to describe the shape of the curve $\theta(\sigma)$ and, using the second derivative, they manage to identify the critical stress and strain ($\sigma_c$ and $\varepsilon_c$).

As introduced earlier, the determination of critical values for the initiation of DRX involves determining the inflection point located on the $\theta(\sigma)$ curve between $\varepsilon=0$ and $\varepsilon_c$.
One of the first difficulties is to numerically evaluate the value of the derivative of stress versus strain, since experimental data from compression tests show oscillations, as illustrated in Figure \ref{fig:RawData}.
It is therefore not possible to calculate the derivative $\theta(\sigma)=\frac{\partial \sigma}{\partial \varepsilon}$ directly.

To get around this problem, some authors recommend smoothing the experimental data (using the Excel solver for example \cite{Najafizadeh-2006}) or identifying an analytical form (e.g. a polynomial of relatively high order) using a minimization algorithm.
But, depending on the degree of the polynomial and for the same stress/strain curve, several solutions can be obtained, leading to the definition of several possible forms for the $\theta(\sigma)$ function.

%----------------------------------------------------------------------------------
\subsubsection{ANN based filtering model's archicture \label{subsec:ANNbasics}}
%----------------------------------------------------------------------------------
In order to improve the results obtained from the critical value identification procedure, we propose here to use the universal approximator capability of artificial neural networks to replace each strain-hardening curve by its representation in the form of an ANN with 1 input (strain) and 1 output (stress).
In such a network, each layer of neurons is connected to the one before it and the one after it by weighted connections.
Any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately preceding layer $(k-1)$, containing $m$ neurons, given by the following~equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i$th neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j$th neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i$th neuron of layer $k$ and the $j$th neuron of layer $(k-1)$, and $b_i\lay{k}$ is the associated bias of the $i$th neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN, which we have to adjust during the training procedure.
For the proposed model, we selected the Sigmoid activation function, so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

No activation function was used for the output neuron of the ANN as usually done in a regression application.
The architecture chosen for our application corresponds to 2 hidden layers of 7 and 5 neurons respectively, as illustrated in Figure \ref{fig:ANN-7-5}.
If $n$ and $m$ are the number of neurons of the first and second hidden layers respectively, the number of internal parameters $N_{int}$ of the model is given by $N_{int}=m(2+n)+2n+1$.
So that for a $7-5$ network, $N_{int}=60$.
\begin{figure}[H]
%\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-7-5}
\caption{Two hidden layers artificial neural network architecture with 1 input neuron (green) and 1 output neuron (red).}
\label{fig:ANN-7-5}
\end{figure}
The Python program used for training the neural network was developed using the specialized Python library, Tensorflow \cite{Abadi-2016}.
The Adaptive Moment Estimation (ADAM) optimizer \cite{Kingma-2015} was used for the training phase.

>>>>

%----------------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%----------------------------------------------------------------------------------

Dynamic recrystallization is a process observed in specific materials, mainly metals and alloys, when undergoing non-reversible plastic deformation.
It is a phenomenon frequently seen in hot working operations, including hot rolling, hot forging, and hot extrusion.
When a metal or alloy undergoes significant plastic deformation at high temperatures, its existing grain structure is disrupted.
During this deformation process, new, smaller grains are formed, which is referred to as dynamic recrystallization (DRX).
DRX differs from static recrystallization (SRX), which takes longer to form new grains and occurs at lower temperatures.
The amount of energy stored in the metal due to plastic deformation is the driving force behind the DRX process.
As the metal deforms, internal energy increases.
At high temperatures, metal atoms have enough mobility to rearrange themselves, forming new equiaxed grains free of strain.
Typically, these newly formed grains are much smaller than the original grains of the undeformed material.
Their growth is accelerated by the stored energy and high temperatures.
DRX provides several benefits in the hot forming process by refining the grain size, enhancing the mechanical properties of the material, including strength and toughness.
Additionally, it decreases the total strain energy, resulting in a more uniform microstructure and better material properties.

During high-rate hot deformation, dislocation density typically rises while a dynamic recovery (DRC) helps to mitigate such increases.
DRX occurs when the dislocation density rises to an extent beyond the DRC’s accommodation.
This is especially true for low and medium stacking fault energies \OP{what is it?} in fcc metals.\OPP{can you explain what you mean?}
Therefore, understanding the critical conditions for dynamic recrystallization (DRX) is important for modeling industrial processes.
However, the extent of DRX is influenced by factors such as material composition, deformation temperature, strain rate, and degree of deformation.
DRX is a crucial phenomenon that must be considered and controlled in industrial processes that use hot working to shape metals and alloys into products.
By optimizing the parameters, engineers can attain the desired microstructures and mechanical properties in the final product.
Various techniques have been proposed by researchers to establish DRX's critical conditions such as metallography.
However, this technique demands abundant sampling prior to and after the critical deformation.
Additionally, the cooling phase entails phase changes from the hot working temperature, which alter the deformed structure, rendering the metallographic analysis complex.
Hence, simpler techniques such as the development of analytical models are required to determine DRX initiation.
In this regard, multiple studies have been carried out aiming to analytically identify the initiation of DRX.
For instance, Barnett \eal \cite{barnett2000predicting} determined the critical strain for the initiation of DRX using the kinetics of static recrystallization (SRX), which was adapted to allow SRX to begin before the end of deformation.
This approach defines the critical stress required for the initiation of SRX during deformation.
According to Gottstein \eal \cite{gottstein2004prediction}, the critical strain for initiating DRX can also be predicted based on a work hardening model of dislocation density.
The onset of dynamic recrystallization (DRX) may be identified through the inflection point in the strain hardening rate $\theta(\sigma)=\frac{\partial \sigma}{\partial \varepsilon}$ where $\sigma$ is the stress and $\varepsilon$ is the strain.
This understanding is crucial for materials processing and design purposes.

The DRX phenomenon, based on the inflection point, was first studied by Ryan and McQueen \cite{ryan1989dynamic, ryan1990dynamic, ryan1990flow}.
They observed the appearance of an inflection point on the $\theta(\sigma)$ \OP{I suggest the following writing: $\theta(\sigma)$ instead of $\theta=f(\sigma)$} curve before the peak stress $\sigma_p$ and attributed the beginning of DRX and the critical conditions to this point.
Later, Poliak and Jonas \cite{Poliak-1996,ei2003initiation,ei2003critical,jonas2003critical} confirmed the hypothesis of DRX describing the inflection point by demonstrating that it is associated with thermodynamic energy released during dislocation movements.
This indicates the onset of an additional softening mechanism, in addition to DRC.
Najafizadeh and Jonas \cite{najafizadeh2006predicting} simplified Poliak and Jonas' thermodynamic model using an analytical technique to determine the critical conditions for initiating DRX.
Their widely-used approach relies on a third-degree polynomial to describe the curve $\theta(\sigma)$ and the second derivative to identify the critical stress $\sigma_c$ and strain $\varepsilon_c$.
Once calculated, the critical conditions serve as input data for the Johnson-Mehl-Avrami-Kolmogorov (JMAK) model \cite{Avrami-1939}.
This model is commonly utilized to estimate the volumetric fraction of dynamic recrystallization $X_{drx}$ during hot working.
Li \eal \cite{li2015experimental} investigated the DRX properties of micro-alloyed plastic molding steel by utilizing the Avrami kinetics model equation and the Estrin and Mecking mathematical model \cite{estrin1984unified,mecking1981kinetics} to determine essential parameters.

@@

The precision of the critical DRX conditions is closely linked to the method of obtaining the curve $\theta(\sigma)$.
Using experimental raw data directly does not yield functional curves, thus smoothing the $\sigma(\varepsilon)$ \OP{I suggest the following writing: $\sigma(\varepsilon)$ instead of $\sigma=f(\varepsilon)$} curve as much as possible or approximating it with a polynomial function that can reproduce the curve and deducing the $\theta(\sigma)$ curve from it is commonly used.
The drawback of both of these methods is that they do not always ensure a satisfactory smoothing of the curve $\sigma(\varepsilon)$.
Nevertheless, the Universal Approximator, Artificial Neural Networks (ANN), has demonstrated effectiveness in various domains, such as predicting the flow behavior of materials.
Hence, this study offers a replacement of the two smoothing methods with an ANN to procure the critical conditions for DRX.
In other words, the proposed method suggests utilizing an artificial neural network (ANN) to directly model the relationship between stress $\sigma$ and rate hardening $\theta$, removing the need for explicit smoothing of experimental data.
By properly training the ANN with relevant data, it can learn the underlying relationship between stress and dynamic recrystallization (DRX), ultimately providing a more precise estimation of the critical conditions for DRX.

@@

The experimental tests used in this work are identical to those previously published by Tize Mha \eal \cite{TizeMha-2023}.
For additional information, readers can refer to the publications by our research group \cite{Pantale-2021, Pantale-2023}.
Nonetheless, we will outline the key aspects necessary to comprehend the suggested methodology.
%The first publication focused on developing a constitutive law capable of predicting material flow, for which a comparative study between 5 analytical models and a neural network-based model was conducted.
%At the conclusion of this comparative study, the Artificial Neural Network (ANN) model proved to be the best performing.
%In this new study, the ANN model will be utilized for predicting dynamic recrystallization, as explained in the introduction.
%It's worth noting that the ANN model identified in the first publication will also be employed for managing the plastic behavior of the material.

%----------------------------------------------------------------------------------
\subsection{Experimental procedure and compression tests results\label{subsec:ExperimentalProcedure}}
%----------------------------------------------------------------------------------

The material utilized for this study is a medium-carbon steel with the chemical composition presented in Table \ref{tab:Composition}.
\begin{table}[H]
\centering
\caption{Chemical composition of medium carbon steel. Fe = balance.}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{LCCCCCCC}
\hline
\textbf{Element} & \textbf{C} & \textbf{Mn} & \textbf{Mo} & \textbf{Si} & \textbf{Ni} & \textbf{Cr} & \textbf{Cu} \\
\hline
\textbf{Wt~\%} & $0.30$ & $0.89$ & $0.52$ & $0.34$ & $0.68$ & $1.86$ & $0.17$ \\
\hline
\end{tabularx}
\label{tab:Composition}
\end{table}
Hot compression tests were conducted on cylinders with an initial diameter of $\phi=10$~mm and a height of $h=15$~mm using a Gleeble-3800 thermomechanical simulator.
The tests were conducted at 5 temperature levels (ranging from $1050\celsius$ to $1250\celsius$) and 6 strain rates (ranging from $0.001~\ps$ to $5~\ps$).
The imposed displacement was set to $d=6$~mm, resulting in a reduction of $40\%$ of the initial height.
To reduce friction during testing, thin tantalum sheets were utilized as a lubricant on the contact surface of the anvils and specimens.
In order to remove thermal gradients, the samples were heated at a speed of $2$\celsius/s until reaching a temperature of $1260\celsius$, then held at that temperature for $5$~minutes.
They were subsequently cooled down to the test temperature at a speed of $1$\celsius/s and left at a consistent temperature for $1$~minute prior to being formed.
During compression, the sample temperature is maintained at a constant level through the machine's thermal control system.
Subsequently, the structure is frozen at a fast pace in order to preserve the microstructure for future analysis.

The raw stress-strain data is exported from the Gleeble thermomechanical simulator as true stress and true strain, and the set of 30 flow stress curves $\sigma$ versus strain $\varepsilon$ obtained from compression tests for each test condition is reported in Figure \ref{fig:RawData}.
All strain/stress data is composed of 701 equidistant strain values from $\varepsilon=0.0$ to $\varepsilon=0.7$ in $0.01$ increments.
\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/rawData}
\caption{Stress--strain curves of medium carbon alloy extracted from the Gleeble device for 5 temperatures ($T$) and 6 strain rates ($\mdot\varepsilon$).}
\label{fig:RawData}
\end{figure}

The flow stress ($\sigma$) increases as the strain rate ($\mdot\varepsilon$) rises but decreases as the temperature ($T$) increases.
Notably, the strain also affects the flow stress.
At the lowest strain rates, the flow stress increases up to $\sigma_p$ as the strain approaches a value of about $\varepsilon_p=0.2$ to $0.3$.
Then, it decreases to maintain a relatively constant value throughout the entire test.
Conversely, at strain rates above $1~\ps$, the flow stress increases steadily throughout the test.
The nominal rise in stress observed at low strain rates with large strain values is attributed to friction between the sample and anvil during testing, as noted by Galos \eal \cite{Galos-2022}.
Furthermore, this frictional effect becomes more apparent as lubrication reduces over time.

The initial increase in stress seen during deformation up to $\varepsilon=0.1$ is believed to be due to work hardening (WH).
After $\varepsilon=0.1$ up to $0.2$, the flow stress indicates a steady decrease as the stress level rises until a peak or inflection point is attained.
This peak suggests that thermal softening is now the predominant influence, overtaking work hardening.
At this stage, the stress-strain curve displays three distinct patterns as strain increases.
The first pattern is a gradual decrease to a steady state with DRV/DRX softening, which is observed at all deformation temperatures and strain rates ranging from $\mdot\varepsilon=0.001~\ps$ to $0.1~\ps$, except at $1050\celsius$ and $1100\celsius$.
The second pattern is higher stress levels without significant softening and work hardening at $1050\celsius$ and $1100\celsius$ with a strain rate of $0.1~\ps$.
The last pattern is a continuous increase with significant work hardening observed for all deformation temperatures and a strain rate above $1~\ps$.
This indicates that DRX softening happens at high temperatures and low strain rates,
and the rate of softening due to DRX slows down at higher strain rates and lower temperatures due to higher work hardening rates.
The result is that both the peak stress $\sigma_p$ and the onset of steady-state flow occur at higher strain levels.
The decrease in stress detected across all temperatures and strain rates of $\mdot\varepsilon=0.001-5.0~\ps$ is attributed to the event of dynamic recrystallization (DRX).

Dynamic recrystallization is a significant phenomenon for regulating a material's microstructural and mechanical properties during hot working.
In some materials like aluminum, DRX can counterbalance work hardening, leading to a rapid emergence of a plateau.
However, in many austenitic steels, the kinetics of restoration are feeble, causing DRX to activate at a critical strain accumulation condition.
Due to DRX's significant impact on the material's flow stress at high temperatures, it unavoidably affects the material's microstructure following hot working.
Therefore, it is critical to accurately predict the start of DRX to effectively characterize the material.
Several factors, such as the chemical composition of the material, the grain size before deformation, the mode of deformation, and the deformation parameters (deformation rate, temperature, and deformation) condition dynamic recrystallization (DRX), which can only be triggered once the critical deformation is reached and can be determined via metallography.
However, this technique necessitates extensive sampling pre- and post-critical deformation.
In addition, the cooling phase alters the deformed structure during the descent from the hot working temperature, resulting in complicating metallographic analysis.
Hence simpler techniques, like developing analytical models have been implemented.
Several studies analyzing the initiation of DRX have been conducted with this objective in mind.

In this paper, we will utilize the Johnson-Mehl-Avrami-Kohnogorov (JMAK) phenomenological model \cite{Avrami-1939} to calculate the development of Dynamic Recrystallization (DRX) of the material during deformation.
The information obtained from the Gleeble thermosimulator will be utilized to identify the material flow law and the parameters of the DRX model.
We obtain the parameters of the JMAK model through a two-step process.
First, critical stresses and strains are determined using the ANN model to filter curves effectively.
Then, these critical values are used as input data to calculate parameters in the JMAK model, with the goal of forecasting the volume fraction of dynamic recrystallization.

@@

In this subsection we will determine the critical deformation required to initiate recrystallization.
Ryan and McQueen \cite{Ryan-1989, Ryan-1990, Ryan-1990-2} were the pioneers in investigating this phenomenon.
Observing the appearance of an inflection point on the curve $\theta(\sigma)=\frac{\partial \sigma}{\partial \varepsilon}$ before the peak stress $\sigma_p$, the start of DRX was attributed to this inflection point, and therefore the critical values of the strain ($\varepsilon_c$) and stress ($\sigma_c$) respectively.
Later, Poliak and Jonas \cite{Poliak-1996, Poliak-2003, Poliak-2003-2, Jonas-2003} demonstrated that this inflection point is associated with a release of thermodynamic free energy during dislocation motion, supporting the DRX hypothesis.
Najafizadeh and Jonas \cite{Najafizadeh-2006} developed an analytical technique to determine the critical conditions for DRX initiation, simplifying the thermodynamic model of Poliak and Jonas.
Their widely-used approach adopts a $3^\text{rd}$ degree polynomial to describe the curve $\theta(\sigma)$, enabling the identification of critical stress and strain ($\sigma_c$ and $\varepsilon_c$) through the second derivative.

As previously stated, critical values for DRX initiation are determined by identifying the inflection point on the $\theta(\sigma)$ curve between $\varepsilon=0$ and $\varepsilon_c$.
One challenge is calculating the derivative of stress versus strain values numerically, as compression test data exhibits oscillations, as depicted in Figure \ref{fig:RawData}.
It is not feasible to compute the derivative $\theta(\sigma)=\frac{\partial \sigma}{\partial \varepsilon}$ directly.

To overcome this difficulty, certain authors propose smoothing out the empirical data (perhaps through the utilization of the Excel solver \cite{Najafizadeh-2006}) or discovering an analytical function (such as a polynomial with a relatively high degree) through a minimization algorithm.
But, depending on the degree of the polynomial and for the same stress/strain curve, multiple solutions may be obtained.
This results in multiple possible forms for the $\theta(\sigma)$ function.

%----------------------------------------------------------------------------------
\subsubsection{ANN based filtering model's archicture \label{subsec:ANNbasics}}
%----------------------------------------------------------------------------------

To enhance the outcomes attained from the critical value identification process, we suggest employing the universal approximator capability of artificial neural networks.
This can be done by characterizing each strain-hardening curve using an ANN with just one input (strain) and one output (stress).
In this type of network, each layer of neurons is linked to the preceding and subsequent layers through weighted connections.
Any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately preceding layer $(k-1)$, containing $m$ neurons, given by the following~equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i$th neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j$th neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i$th neuron of layer $k$ and the $j$th neuron of layer $(k-1)$, and $b_i\lay{k}$ is the associated bias of the $i$th neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN, which we have to adjust during the training procedure.
For the proposed model, we selected the Sigmoid activation function, so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

No activation function was used for the output neuron of the ANN, as is typically omitted in regression applications.
The chosen architecture for our application consists of two hidden layers, one with 7 neurons and the other with 5 neurons, as illustrated in Figure \ref{fig:ANN-7-5}.
To calculate the number of internal parameters $N_{int}$ in the model, where $n$ and $m$ represent the number of neurons in the first and second hidden layers, respectively, the following formula may be used: $N_{int}=m(2+n)+2n+1$.
So that for a $7-5$ network, $N_{int}=60$.
\begin{figure}[H]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-7-5}
\caption{Two hidden layers artificial neural network architecture with 1 input neuron (green) and 1 output neuron (red).}
\label{fig:ANN-7-5}
\end{figure}
The Python program utilized for training the neural network was created with the dedicated Python library, Tensorflow \cite{Abadi-2016}.
The training phase employed the Adaptive Moment Estimation (ADAM) optimizer \cite{Kingma-2015}.

%----------------------------------------------------------------------------------
\subsubsection{ANN filtering based model's application \label{subsec:ANNapplication}}
%----------------------------------------------------------------------------------

Since the dynamic recrystallization phenomenon only occurs for strain rates between $0.001~\ps$ and $1.0~\ps$, we will use only the first 20 stress/strain curves to identify the DRX critical parameters.
Each of these 20 curves consists of 701 pairs of strain-stress values and was independently processed by a neural network, resulting in the identification of 20 networks.
As an example, Figure \ref{fig:AnnFit} displays the stress/strain curve experimental values for $\mdot\varepsilon=0.001~\ps$ and $T=1050\celsius$ (in blue) and the neural network results identified from this data (in red).
